---
title: "Armed Services Committees NOMINATE Visualization"
output: github_document
---

## Armed Services Committee Rosters
# The membership for the House and Senate Armed Services Committees (including the House Committee on National Security in the 104th and 105th Congresses) was derived from the committees' reports of activities.
# Committees published these reports on varying schedules, sometimes long after the Congress to which they referred. The reports were unavailable for some Congresses, and I assembled membership information from other sources, including C-SPAN hearings, which often list the committee membership. The C-SPAN website lists the membership for the committees by Congress but without amplifying data such as information on members who left or joined the committee during the Congress.
# These lists include ALL members of the committee during that Congress, including those who left the committee or joined late. There are a few exceptions, for instance in the 102nd Congress, I included Mary Bono but not Sonny Bono because he died at the beginning of the session and played no part in the hearings covered by this project.
# There are several issues with NOMINATE scores, but they are the most readily available and consistent measure of ideology.
# 1. NOMINATE scores are consistent for a single member across their membership in a legislative body, so they do not capture ideological evolution. Therefore, these measures may under- or overstate a member's ideological extremism at any given point. 
# 2. The NOMINATE dimensions do not directly address military and national security issues. It is possible that there are outliers who score extremely conservative on the economic or racial scale by are liberal on national security issues. However, that seems less likely during the period of this study--for instance Democrats like Joe Lieberman who were hawkish and conservative on national security were also among the most conservative members of their party on other issues. 

## Code

# Load required libraries

```{r, echo = FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(haven)
library(here)           # sets directory path
library(ggplot2)        # plotting functionality including attr for labels
library(knitr)          # enables running code from RMD file
library(here)           # could not get this to work with Proton Drive
library(RColorBrewer)   # RColorBrewer provides color palettes for plots
```

# Set home folder
```{r, echo = FALSE}
# here() starts at /Library/CloudStorage/ProtonDrive-james.vizzard@proton.me-folder/James/201801-20YYMM_PhD/P_Dissertation
i_am("/Documents/Dissertation/")
```

# Load data and create subsets by chamber by congress
```{r, echo = FALSE}
# using the working csv files constructed from the base VoteView data with the addition of armed services committee membership, create data frames for each chamber and armed services committee in each congress.

house <- read_csv("NOMINATE_House_202506_WORKING.csv")

# set labels
attr(house$congress, "label") <- "Congress"
attr(house$chamber, "label") <- "Chamber"
attr(house$icpsr, "label") <- "ICPSR Identifier"
attr(house$state_icpsr, "label") <- "ICPSR State Identifier"
attr(house$district_code, "label") <- "District Code"
attr(house$state_abbrev, "label") <- "State"
attr(house$party_code, "label") <- "Party Code"
attr(house$party, "label") <- "Party"
attr(house$occupancy, "label") <- "ICPSR Occupancy Code"
attr(house$last_means, "label") <- "Means of Attaining Office"
attr(house$bioname, "label") <- "Member Name"
attr(house$bioguide_id, "label") <- "Biographical Directory of Congress ID"
attr(house$born, "label") <- "Date of Birth"
attr(house$died, "label") <- "Date of Death"
attr(house$nominate_dim1, "label") <- "NOMINATE Dim1"
attr(house$nominate_dim2, "label") <- "NOMINATE Dim2"
attr(house$nominate_log_likelihood, "label") <- "Log-Likelihood of the NOMINATE Estimate"
attr(house$nominate_geo_mean_probability, "label") <- "Geometric Mean Probability" 
attr(house$nominate_number_of_votes, "label") <- "Votes Cast by Member in Given Congress"
attr(house$nokken_poole_dim1, "label") <- "Nokken-Poole First Dimension Estimate"
attr(house$nokken_poole_dim2, "label") <- "Nokken-Poole Second Dimension Estimate"
attr(house$hasc, "label") <- "HASC Member"

house102 <- subset( x = house,           # data frame is house
              subset = congress == 102   # keep only 102nd Congress
                  )
hasc102 <- subset( x = house102,         # data frame is house102
                   subset = hasc == 1.   # keep only HASC members
                   )
house103 <- subset( x = house,           # repeat for 103-114 Congresses
              subset = congress == 103   
                  )
hasc103 <- subset( x = house103,
                   subset = hasc == 1
                   )
house104 <- subset( x = house,          
              subset = congress == 104   
                  )
hasc104 <- subset( x = house104,
                   subset = hasc == 1
                   )
house105 <- subset( x = house,           
              subset = congress == 105   
                  )
hasc105 <- subset( x = house105,
                   subset = hasc == 1
                   )
house106 <- subset( x = house,          
              subset = congress == 106   
                  )
hasc106 <- subset( x = house106,
                   subset = hasc == 1
                   )
house107 <- subset( x = house,         
              subset = congress == 107  
                  )
hasc107 <- subset( x = house107,
                   subset = hasc == 1
                   )
house108 <- subset( x = house,          
              subset = congress == 108   
                  )
hasc108 <- subset( x = house108,
                   subset = hasc == 1
                   )
house109 <- subset( x = house,           
              subset = congress == 109   
                  )
hasc109 <- subset( x = house109,
                   subset = hasc == 1
                   )
house110 <- subset( x = house,           
              subset = congress == 110   
                  )
hasc110 <- subset( x = house110,
                   subset = hasc == 1
                   )
house111 <- subset( x = house,           
              subset = congress == 111   
                  )
hasc111 <- subset( x = house111,
                   subset = hasc == 1
                   )
house112 <- subset( x = house,           
              subset = congress == 112   
                  )
hasc112 <- subset( x = house112,
                   subset = hasc == 1
                   )
house113 <- subset( x = house,           
              subset = congress == 113   
                  )
hasc113 <- subset( x = house113,
                   subset = hasc == 1
                   )
house114 <- subset( x = house,           
              subset = congress == 114   
                  )
hasc114 <- subset( x = house114,
                   subset = hasc == 1
                   )
```

```{r}
senate <- read_csv("NOMINATE_Senate_202506_WORKING.csv")

# set labels
attr(senate$congress, "label") <- "Congress"
attr(senate$chamber, "label") <- "Chamber"
attr(senate$icpsr, "label") <- "ICPSR Identifier"
attr(senate$state_icpsr, "label") <- "ICPSR State Identifier"
attr(senate$district_code, "label") <- "District Code"
attr(senate$state_abbrev, "label") <- "State"
attr(senate$party_code, "label") <- "Party Code"
attr(senate$party, "label") <- "Party"
attr(senate$occupancy, "label") <- "ICPSR Occupancy Code"
attr(senate$last_means, "label") <- "Means of Attaining Office"
attr(senate$bioname, "label") <- "Member Name"
attr(senate$bioguide_id, "label") <- "Biographical Directory of Congress ID"
attr(senate$born, "label") <- "Date of Birth"
attr(senate$died, "label") <- "Date of Death"
attr(senate$nominate_dim1, "label") <- "NOMINATE Dim1"
attr(senate$nominate_dim2, "label") <- "NOMINATE Dim2"
attr(senate$nominate_log_likelihood, "label") <- "Log-Likelihood of the NOMINATE Estimate"
attr(senate$nominate_geo_mean_probability, "label") <- "Geometric Mean Probability" 
attr(senate$nominate_number_of_votes, "label") <- "Votes Cast by Member in Given Congress"
attr(senate$nokken_poole_dim1, "label") <- "Nokken-Poole First Dimension Estimate"
attr(senate$nokken_poole_dim2, "label") <- "Nokken-Poole Second Dimension Estimate"
attr(senate$sasc, "label") <- "SASC Member"

sen102 <- subset( x = senate,            # data frame is senate
              subset = congress == 102   # keep only 102nd Congress
                  )
sasc102 <- subset( x = sen102,           # data frame is sen102
                   subset = sasc == 1.   # keep only SASC members
                   )
sen103 <- subset( x = senate,           # repeat for 103-114 Congresses
              subset = congress == 103   
                  )
sasc103 <- subset( x = sen103,
                   subset = sasc == 1
                   )
sen104 <- subset( x = senate,           
              subset = congress == 104   
                  )
sasc104 <- subset( x = sen104,
                   subset = sasc == 1
                   )
sen105 <- subset( x = senate,           
              subset = congress == 105   
                  )
sasc105 <- subset( x = sen105,
                   subset = sasc == 1
                   )
sen106 <- subset( x = senate,           
              subset = congress == 106   
                  )
sasc106 <- subset( x = sen106,
                   subset = sasc == 1
                   )
sen107 <- subset( x = senate,           
              subset = congress == 107   
                  )
sasc107 <- subset( x = sen107,
                   subset = sasc == 1
                   )
sen108 <- subset( x = senate,           
              subset = congress == 108   
                  )
sasc108 <- subset( x = sen108,
                   subset = sasc == 1
                   )
sen109 <- subset( x = senate,           
              subset = congress == 109   
                  )
sasc109 <- subset( x = sen109,
                   subset = sasc == 1
                   )
sen110 <- subset( x = senate,           
              subset = congress == 110   
                  )
sasc110 <- subset( x = sen110,
                   subset = sasc == 1
                   )
sen111 <- subset( x = senate,           
              subset = congress == 111   
                  )
sasc111 <- subset( x = sen111,
                   subset = sasc == 1
                   )
sen112 <- subset( x = senate,           
              subset = congress == 112   
                  )
sasc112 <- subset( x = sen112,
                   subset = sasc == 1
                   )
sen113 <- subset( x = senate,           
              subset = congress == 113   
                  )
sasc113 <- subset( x = sen113,
                   subset = sasc == 1
                   )
sen114 <- subset( x = senate,           
              subset = congress == 114   
                  )
sasc114 <- subset( x = sen114,
                   subset = sasc == 1
                   )
```

Calculate summary statistics for each house and each oversight committee.

```{r echo = FALSE}

summ_102_hr <-house102 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_102_hr, file = "summ_102_hr.csv",
          row.names = FALSE)

summ_102_hasc <- hasc102 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_102_hasc, file = "summ_102_hasc.csv",
          row.names = FALSE)

summ_102_sen <-sen102 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_102_sen, file = "summ_102_sen.csv",
          row.names = FALSE)

summ_102_sasc <- sasc102 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_102_sasc, file = "summ_102_sasc.csv",
          row.names = FALSE)

summ_103_hr <-house103 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_103_hr, file = "summ_103_hr.csv",
          row.names = FALSE)

summ_103_hasc <- hasc103 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_103_hasc, file = "summ_103_hasc.csv",
          row.names = FALSE)

summ_103_sen <-sen103 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_103_sen, file = "summ_103_sen.csv",
          row.names = FALSE)

summ_103_sasc <- sasc103 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_103_sasc, file = "summ_103_sasc.csv",
          row.names = FALSE)

summ_104_hr <-house104 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_104_hr, file = "summ_104_hr.csv",
          row.names = FALSE)

summ_104_hasc <- hasc104 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_104_hasc, file = "summ_104_hasc.csv",
          row.names = FALSE)

summ_104_sen <-sen104 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_104_sen, file = "summ_104_sen.csv",
          row.names = FALSE)

summ_104_sasc <- sasc104 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_104_sasc, file = "summ_104_sasc.csv",
          row.names = FALSE)

summ_105_hr <-house105 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_105_hr, file = "summ_105_hr.csv",
          row.names = FALSE)

summ_105_hasc <- hasc105 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_105_hasc, file = "summ_105_hasc.csv",
          row.names = FALSE)

summ_105_sen <-sen105 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_105_sen, file = "summ_105_sen.csv",
          row.names = FALSE)

summ_105_sasc <- sasc105 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_105_sasc, file = "summ_105_sasc.csv",
          row.names = FALSE)

summ_106_hr <-house106 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_106_hr, file = "summ_106_hr.csv",
          row.names = FALSE)

summ_106_hasc <- hasc106 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_106_hasc, file = "summ_106_hasc.csv",
          row.names = FALSE)

summ_106_sen <-sen106 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_106_sen, file = "summ_106_sen.csv",
          row.names = FALSE)

summ_106_sasc <- sasc106 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_106_sasc, file = "summ_106_sasc.csv",
          row.names = FALSE)

summ_107_hr <-house107 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_107_hr, file = "summ_107_hr.csv",
          row.names = FALSE)

summ_107_hasc <- hasc107 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_107_hasc, file = "summ_107_hasc.csv",
          row.names = FALSE)

summ_107_sen <-sen107 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_107_sen, file = "summ_107_sen.csv",
          row.names = FALSE)

summ_107_sasc <- sasc107 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_107_sasc, file = "summ_107_sasc.csv",
          row.names = FALSE)

summ_108_hr <-house108 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_108_hr, file = "summ_108_hr.csv",
          row.names = FALSE)

summ_108_hasc <- hasc108 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_108_hasc, file = "summ_108_hasc.csv",
          row.names = FALSE)

summ_108_sen <-sen108 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_108_sen, file = "summ_108_sen.csv",
          row.names = FALSE)

summ_108_sasc <- sasc108 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_108_sasc, file = "summ_108_sasc.csv",
          row.names = FALSE)

summ_109_hr <-house109 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_109_hr, file = "summ_109_hr.csv",
          row.names = FALSE)

summ_109_hasc <- hasc109 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_109_hasc, file = "summ_109_hasc.csv",
          row.names = FALSE)

summ_109_sen <-sen109 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_109_sen, file = "summ_109_sen.csv",
          row.names = FALSE)

summ_109_sasc <- sasc109 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_109_sasc, file = "summ_109_sasc.csv",
          row.names = FALSE)

summ_110_hr <-house110 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_110_hr, file = "summ_110_hr.csv",
          row.names = FALSE)

summ_110_hasc <- hasc110 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_110_hasc, file = "summ_110_hasc.csv",
          row.names = FALSE)

summ_110_sen <-sen110 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_110_sen, file = "summ_110_sen.csv",
          row.names = FALSE)

summ_110_sasc <- sasc110 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_110_sasc, file = "summ_110_sasc.csv",
          row.names = FALSE)

summ_111_hr <-house111 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_111_hr, file = "summ_111_hr.csv",
          row.names = FALSE)

summ_111_hasc <- hasc111 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_111_hasc, file = "summ_111_hasc.csv",
          row.names = FALSE)

summ_111_sen <-sen111 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_111_sen, file = "summ_111_sen.csv",
          row.names = FALSE)

summ_111_sasc <- sasc111 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_111_sasc, file = "summ_111_sasc.csv",
          row.names = FALSE)

summ_112_hr <-house112 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_112_hr, file = "summ_112_hr.csv",
          row.names = FALSE)

summ_112_hasc <- hasc112 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_112_hasc, file = "summ_112_hasc.csv",
          row.names = FALSE)

summ_112_sen <-sen112 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_112_sen, file = "summ_112_sen.csv",
          row.names = FALSE)

summ_112_sasc <- sasc112 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_112_sasc, file = "summ_112_sasc.csv",
          row.names = FALSE)

summ_113_hr <-house113 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_113_hr, file = "summ_113_hr.csv",
          row.names = FALSE)

summ_113_hasc <- hasc113 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_113_hasc, file = "summ_113_hasc.csv",
          row.names = FALSE)

summ_113_sen <-sen113 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_113_sen, file = "summ_113_sen.csv",
          row.names = FALSE)

summ_113_sasc <- sasc113 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_113_sasc, file = "summ_113_sasc.csv",
          row.names = FALSE)

summ_114_hr <-house114 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_114_hr, file = "summ_114_hr.csv",
          row.names = FALSE)

summ_114_hasc <- hasc114 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_114_hasc, file = "summ_114_hasc.csv",
          row.names = FALSE)

summ_114_sen <-sen114 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_114_sen, file = "summ_114_sen.csv",
          row.names = FALSE)

summ_114_sasc <- sasc114 %>%
  group_by(party) %>% 
  summarize(min = min(nominate_dim1),
            median = median(nominate_dim1),
            mean = mean(nominate_dim1),
            max = max(nominate_dim1))
write.csv(summ_114_sasc, file = "summ_114_sasc.csv",
          row.names = FALSE)



```







## Plots of committee and chamber NOMINATE scores
```{r fig.width=6, fig.height=3.75, echo = FALSE}

# plots of HASC and the House of Representatives in 102nd Congress

# set default colors
# set colors for plots
party_colors <- c("blue", "red", "purple") 

# scatter plot
phasc102 <- ggplot(data = hasc102,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc102 + geom_point() + 
  geom_smooth(method = lm) +
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_102.pdf")
ggsave(filename = "pointplot_hasc_102.png")

# boxplot HASC
bhasc102 <- ggplot(data = hasc102,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc102 + geom_boxplot() +
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_102.png")
ggsave(filename = "boxplot_hasc_102.pdf")

# Scatterplot House of Representatives
phr102 <- ggplot(data = house102,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr102 + geom_point() + 
  geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_102.png")
ggsave(filename = "pointplot_hr_102.pdf")

# boxplot House of Representatives
bhr102 <- ggplot(data = house102,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhr102 + geom_boxplot() +
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hr_102.png")
ggsave(filename = "boxplot_hr_102.pdf")

# 103rd Congress
# scatterplot HASC
phasc103 <- ggplot(data = hasc103,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc103 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 103rd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_103.png")
ggsave(filename = "pointplot_hasc_103.pdf")

# boxplot HASC
bhasc103 <- ggplot(data = hasc103,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc103 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 103rd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_103.png")
ggsave(filename = "boxplot_hasc_103.pdf")

# Scatterplot House of Representatives in
phr103 <- ggplot(data = house103,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr103 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 103rd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_103.png")
ggsave(filename = "pointplot_hr_103.pdf")

# 104th Congress
# scatterplot HASC
phasc104 <- ggplot(data = hasc104,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc104 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 104th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_104.png")
ggsave(filename = "pointplot_hasc_104.pdf")

# boxplot HASC
bhasc104 <- ggplot(data = hasc104,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc104 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 104th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_104.png")
ggsave(filename = "boxplot_hasc_104.pdf")

# Scatterplot House of Representatives
phr104 <- ggplot(data = house104,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr104 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 104th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_104.png")
ggsave(filename = "pointplot_hr_104.pdf")

# 105th Congress
# scatterplot HASC
phasc105 <- ggplot(data = hasc105,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc105 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 105th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_105.png")
ggsave(filename = "pointplot_hasc_105.pdf")

# boxplot HASC
bhasc105 <- ggplot(data = hasc105,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc105 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 105th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_105.png")
ggsave(filename = "boxplot_hasc_105.pdf")

# Scatterplot House of Representatives
phr105 <- ggplot(data = house105,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr105 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 105th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_105.png")
ggsave(filename = "pointplot_hr_105.pdf")

# 106th Congress
# scatterplot HASC
phasc106 <- ggplot(data = hasc106,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc106 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 106th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_106.png")
ggsave(filename = "pointplot_hasc_106.pdf")

# boxplot HASC
bhasc106 <- ggplot(data = hasc106,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc106 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 106th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_106.png")
ggsave(filename = "boxplot_hasc_106.pdf")

# Scatterplot House of Representatives
phr106 <- ggplot(data = house106,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr106 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 106th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_106.png")
ggsave(filename = "pointplot_hr_106.pdf")

# 107th Congress
# scatterplot HASC
phasc107 <- ggplot(data = hasc107,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc107 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 107th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_107.png")
ggsave(filename = "pointplot_hasc_107.pdf")

# boxplot HASC
bhasc107 <- ggplot(data = hasc107,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc107 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 107th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_107.png")
ggsave(filename = "boxplot_hasc_107.pdf")

# Scatterplot House of Representatives
phr107 <- ggplot(data = house107,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr107 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 107th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_107.png")
ggsave(filename = "pointplot_hr_107.pdf")

# 108th Congress
# scatterplot HASC
phasc108 <- ggplot(data = hasc108,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc108 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 108th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_108.png")
ggsave(filename = "pointplot_hasc_108.pdf")

# boxplot HASC
bhasc108 <- ggplot(data = hasc108,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc108 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 108th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_108.png")
ggsave(filename = "boxplot_hasc_108.pdf")

# Scatterplot House of Representatives
phr108 <- ggplot(data = house108,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr108 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 108th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_108.png")
ggsave(filename = "pointplot_hr_108.pdf")

# 109th Congress
# scatterplot HASC
phasc109 <- ggplot(data = hasc109,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc109 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 109th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_109.png")
ggsave(filename = "pointplot_hasc_109.pdf")

# boxplot HASC
bhasc109 <- ggplot(data = hasc109,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc109 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 109th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_109.png")
ggsave(filename = "boxplot_hasc_109.pdf")

# Scatterplot of House of Representatives
phr109 <- ggplot(data = house109,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr109 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 109th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_109.png")
ggsave(filename = "pointplot_hr_109.pdf")

# 110th Congress
# scatterplot HASC
phasc110 <- ggplot(data = hasc110,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc110 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 110th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_110.png")
ggsave(filename = "pointplot_hasc_110.pdf")

# boxplot HASC
bhasc110 <- ggplot(data = hasc110,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc110 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 110th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_110.png")
ggsave(filename = "boxplot_hasc_110.pdf")

# Scatterplot House of Representatives
phr110 <- ggplot(data = house110,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr110 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 110th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_110.png")
ggsave(filename = "pointplot_hr_110.pdf")

# 111th Congress
# scatterplot HASC
phasc111 <- ggplot(data = hasc111,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc111 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 111th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_111.png")
ggsave(filename = "pointplot_hasc_111.pdf")

# boxplot HASC
bhasc111 <- ggplot(data = hasc111,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc111 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 111th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_111.png")
ggsave(filename = "boxplot_hasc_111.pdf")

# Scatterplot House of Representatives
phr111 <- ggplot(data = house111,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr111 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 111th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_111.png")
ggsave(filename = "pointplot_hr_111.pdf")

# 112th Congress
# scatterplot HASC
phasc112 <- ggplot(data = hasc112,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc112 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 112th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_112.png")
ggsave(filename = "pointplot_hasc_112.pdf")

# boxplot HASC
bhasc112 <- ggplot(data = hasc112,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc112 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 112th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_112.png")
ggsave(filename = "boxplot_hasc_112.pdf")

# Scatterplot House of Representatives
phr112 <- ggplot(data = house112,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr112 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 112th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_112.png")
ggsave(filename = "pointplot_hr_112.pdf")

# 113th Congress
# scatterplot HASC
phasc113 <- ggplot(data = hasc113,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc113 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 113th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_113.png")
ggsave(filename = "pointplot_hasc_113.pdf")

# boxplot HASC
bhasc113 <- ggplot(data = hasc113,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc113 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 113th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_113.png")
ggsave(filename = "boxplot_hasc_113.pdf")

# Scatterplot House of Representatives 
phr113 <- ggplot(data = house113,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr113 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 113th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_113.png")
ggsave(filename = "pointplot_hr_113.pdf")

# 114th Congress
# scatterplot HASC
phasc114 <- ggplot(data = hasc114,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phasc114 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "HASC, 114th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hasc_114.png")
ggsave(filename = "pointplot_hasc_114.pdf")

# boxplot HASC
bhasc114 <- ggplot(data = hasc114,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhasc114 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "HASC, 114th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_hasc_114.png")
ggsave(filename = "boxplot_hasc_114.pdf")

# Scatterplot House of Representatives
phr114 <- ggplot(data = house114,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
phr114 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-1, 1), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "House of Representatives, 114th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_hr_114.png")
ggsave(filename = "pointplot_hr_114.pdf")

# plots of SASC and the Senate
#102nd Congress
# scatter plot SASC
psasc102 <- ggplot(data = sasc102,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc102 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_102.png")
ggsave(filename = "pointplot_sasc_102.pdf")

# boxplot SASC
bsasc102 <- ggplot(data = sasc102,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc102 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_102.png")
ggsave(filename = "boxplot_sasc_102.pdf")

# Scatterplot Senate
psen102 <- ggplot(data = sen102,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen102 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_102.png")
ggsave(filename = "pointplot_sen_102.pdf")

# boxplot Senate
bsen102 <- ggplot(data = sen102,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bhr102 + geom_boxplot() +
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "Senate, 102nd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sen_102.png")
ggsave(filename = "boxplot_sen_102.pdf")

# 103rd Congress
# scatterplot SASC
psasc103 <- ggplot(data = sasc103,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc103 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 103rd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_103.png")
ggsave(filename = "pointplot_sasc_103.pdf")

# boxplot SASC
bsasc103 <- ggplot(data = sasc103,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc103 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 103rd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_103.png")
ggsave(filename = "boxplot_sasc_103.pdf")

# Scatterplot Senate
psen103 <- ggplot(data = sen103,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen103 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 103rd Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_103.png")
ggsave(filename = "pointplot_sen_103.pdf")

# 104th Congress
# scatterplot SASC
psasc104 <- ggplot(data = sasc104,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc104 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 104th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_104.png")
ggsave(filename = "pointplot_sasc_104.pdf")

# boxplot SASC
bsasc104 <- ggplot(data = hasc104,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc104 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 104th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_104.png")
ggsave(filename = "boxplot_sasc_104.pdf")

# Scatterplot Senate
psen104 <- ggplot(data = sen104,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen104 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 104th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_104.png")
ggsave(filename = "pointplot_sen_104.pdf")

# 105th Congress
# scatterplot SASC
psasc105 <- ggplot(data = sasc105,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc105 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 105th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_105.png")
ggsave(filename = "pointplot_sasc_105.pdf")

# boxplot SASC
bsasc105 <- ggplot(data = sasc105,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc105 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
  labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 105th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_105.png")
ggsave(filename = "boxplot_sasc_105.pdf")

# Scatterplot Senate
psen105 <- ggplot(data = sen105,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen105 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 105th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_105.png")
ggsave(filename = "pointplot_sen_105.pdf")

# 106th Congress
# scatterplot SASC
psasc106 <- ggplot(data = sasc106,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc106 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "sASC, 106th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_106.png")
ggsave(filename = "pointplot_sasc_106.pdf")

# boxplot sASC
bsasc106 <- ggplot(data = sasc106,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc106 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 106th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_106.png")
ggsave(filename = "boxplot_sasc_106.pdf")

# Scatterplot Sneate
psen106 <- ggplot(data = sen106,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen106 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 106th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_106.png")
ggsave(filename = "pointplot_sen_106.pdf")

# 107th Congress
# scatterplot SASC
psasc107 <- ggplot(data = sasc107,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc107 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 107th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_107.png")
ggsave(filename = "pointplot_sasc_107.pdf")

# boxplot SASC
bsasc107 <- ggplot(data = sasc107,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc107 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 107th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_107.png")
ggsave(filename = "boxplot_sasc_107.pdf")

# Scatterplot Senate
psen107 <- ggplot(data = sen107,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen107 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 107th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_107.png")
ggsave(filename = "pointplot_sen_107.pdf")

# 108th Congress
# scatterplot SASC
psasc108 <- ggplot(data = sasc108,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc108 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 108th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_108.png")
ggsave(filename = "pointplot_sasc_108.pdf")

# boxplot SASC
bsasc108 <- ggplot(data = sasc108,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc108 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 108th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_108.png")
ggsave(filename = "boxplot_sasc_108.pdf")

# Scatterplot Senate
psen108 <- ggplot(data = sen108,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen108 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 108th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_108.png")
ggsave(filename = "pointplot_sen_108.pdf")

# 109th Congress
# scatterplot SASC
psasc109 <- ggplot(data = sasc109,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc109 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 109th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_109.png")
ggsave(filename = "pointplot_sasc_109.pdf")

# boxplot SASC
bsasc109 <- ggplot(data = sasc109,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc109 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 109th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_109.png")
ggsave(filename = "boxplot_sasc_109.pdf")

# Scatterplot of Senate
psen109 <- ggplot(data = sen109,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen109 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 109th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_109.png")
ggsave(filename = "pointplot_sen_109.pdf")

# 110th Congress
# scatterplot SASC
psasc110 <- ggplot(data = sasc110,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc110 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 110th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_110.png")
ggsave(filename = "pointplot_sasc_110.pdf")

# boxplot SASC
bsasc110 <- ggplot(data = sasc110,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc110 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 110th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_110.png")
ggsave(filename = "boxplot_sasc_110.pdf")

# Scatterplot Senate
psen110 <- ggplot(data = sen110,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen110 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 110th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_110.png")
ggsave(filename = "pointplot_sen_110.pdf")

# 111th Congress
# scatterplot SASC
psasc111 <- ggplot(data = sasc111,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc111 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 111th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_111.png")
ggsave(filename = "pointplot_sasc_111.pdf")

# boxplot SASC
bsasc111 <- ggplot(data = sasc111,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc111 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 111th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_111.png")
ggsave(filename = "boxplot_sasc_111.pdf")

# Scatterplot Senate
psen111 <- ggplot(data = sen111,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen111 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 111th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_111.png")
ggsave(filename = "pointplot_sen_111.pdf")

# 112th Congress
# scatterplot SASC
psasc112 <- ggplot(data = sasc112,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc112 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 112th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_112.png")
ggsave(filename = "pointplot_sasc_112.pdf")

# boxplot SASC
bsasc112 <- ggplot(data = sasc112,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc112 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 112th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_112.png")
ggsave(filename = "boxplot_sasc_112.pdf")

# Scatterplot Senate
psen112 <- ggplot(data = sen112,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen112 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 112th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_112.png")
ggsave(filename = "pointplot_sen_112.pdf")

# 113th Congress
# scatterplot SASC
psasc113 <- ggplot(data = sasc113,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc113 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 113th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_113.png")
ggsave(filename = "pointplot_sasc_113.pdf")

# boxplot SASC NEED TO FIX THIS FOR COLORS
bsasc113 <- ggplot(data = sasc113,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc113 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 113th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_113.png")
ggsave(filename = "boxplot_sasc_113.pdf")

# Scatterplot Senate
psen113 <- ggplot(data = sen113,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen113 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 113th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_113.png")
ggsave(filename = "pointplot_sen_113.pdf")

# 114th Congress
# scatterplot SASC
psasc114 <- ggplot(data = sasc114,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psasc114 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "SASC, 114th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sasc_114.png")
ggsave(filename = "pointplot_sasc_114.pdf")

# boxplot SASC
bsasc114 <- ggplot(data = sasc114,
            mapping = aes(x = party, y = nominate_dim1, color = party))
bsasc114 + geom_boxplot() +
  scale_color_manual(values=party_colors) +
labs(x = "Party", y = "Liberal / Conservative (econ)",
     title = "Ideological Distribution",
     subtitle = "SASC, 114th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "boxplot_sasc_114.png")
ggsave(filename = "boxplot_sasc_114.pdf")

# Scatterplot Senate
psen114 <- ggplot(data = sen114,
            mapping = aes(x = nominate_dim1, y = nominate_dim2, color = party))
psen114 + geom_point() + geom_smooth(method = lm) + 
  scale_color_manual(values = c("D" = "blue", 
                                "R" = "red", 
                                "I" = "purple")) +
  scale_x_continuous(limits=c(-.75,.75), breaks=seq(-.75,.75,.25)) +
  labs(x = "Liberal / Conservative (econ)", y = "Intraparty",
     title = "Ideological Distribution",
     subtitle = "Senate, 114th Congress",
     caption = "Source: DW-NOMINATE")
ggsave(filename = "pointplot_sen_114.png")
ggsave(filename = "pointplot_sen_114.pdf")

```

## Chamber Polarization

## here's where I need to insert the code to calculate the polarization graphics.

---
title: Polarization in Congress
author: Jeff Lewis
description: Provides updated versions the classic Voteview plots of party mean DW-NOMINATE scores of overtime.  Figures and the underlying data are available for download. 
original_date: "2018-02-01" # Set the original date of your contribution in YYYY-MM-DD format here.
date: "`r format(Sys.time(), '%B %d, %Y')`" # Do not modify this
update_delta: 31
tags:
  - blog # Valid tags are: blog, help, data. You may use multiple tags for a post
  - data
---

```{r echo=FALSE, message=FALSE, error=FALSE}
# We use pacman to load and install dependencies.
library(pacman)

# Please list your article's dependencies here, comma separated:
p_load(tidyverse, ggplot2, ggrepel)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we create the polarization data and key graphics the Keith shows and describes in greater detail [here
](https://legacy.voteview.com/political_polarization_2015.htm).  You can download the `Rmarkdown` code that produces this file and all of the data sets and graphics presented [here](party_polarization.Rmd).

This polarization data set includes the following fields (again following Keith):

```
 0.  Chamber
 1.  Congress Number
 2.  First Year of the Congress
 3.  Difference in Party Means - first dimension
 4.  Proportion Moderates
 5.  Proportion of moderate Democrats (-0.25 to +0.25)
 6.  Proportion of moderate Republicans (-0.25 to +0.25)
 7.  Overlap
 8.  Chamber Mean - first dimension
 9.  Chamber Mean - second dimension
10.  Democratic Party Mean - first dimension
11.  Democratic Party Mean - second dimension
12.  Republican Party Mean - first dimension
13.  Republican Party Mean - second dimension
18.  Armed Services Republicans - first dimension
20.  Armed Services Democrats - first dimension
```
# Load data from the polarization csv I file I previously created using the procedures from https://voteview.com/articles/party_polarization
```{r}
 
 polar_dat <- read_csv ("voteview_polarization_data.csv")
 
```

Here are the first few lines of the resulting dataset:

```{r}
head(polar_dat)
write_csv(polar_dat,path="voteview_polarization_data.csv")
```

## Party means on the liberal-conservative dimension over time by chamber

Here are the party means over time plots for each chamber:
```{r echo=FALSE}
polar_dat_long <- polar_dat %>% gather(score,value,-chamber,-year,-congress)
labels <- c("dem.mean.d1"="All D",
            "rep.mean.d1"="All R",
            "ascr"="Armed Svcs R",
            "ascd"="Armed Svcs D")

polarized_plot <- function(chamb) {
  pdatl <- polar_dat_long %>% 
                filter(chamber==chamb,
                       score %in% c("dem.mean.d1","rep.mean.d1",
                                    "north.dem.mean.d1","south.dem.mean.d1")) %>%
                mutate(party=labels[score]) %>%
                ungroup()
  
  gg <- ggplot(data=pdatl,
               aes(x=year,y=value,group=party,col=party)) +
               scale_x_continuous(expand = c(0.15, 0), 
                                 breaks=seq(1991, max(pdatl$year), by=8)) +
               geom_line() + geom_point(size=0.7) + 
               xlab("Year") + ylab("Liberal / Conservative (econ)") + 
               geom_text_repel(data=pdatl %>%
                    filter(year == min(year)),
                  aes(label = party, color = party),
                  size = 3,
                  nudge_x = -8,
                  point.padding = 0.1,
                  segment.color = NA,
                  show.legend = FALSE) +
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              scale_color_manual(values=c("All R"="red","All D"="blue",
                                         "Armed Svcs R"="violet",
                                         "Armed Svcs D"="purple"),guide="none") +
              theme_bw()
  gg
}
```

### House
```{r echo=FALSE}
gg <- polarized_plot("House")
ggsave("voteview_house_party_means.png", plot=gg, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_house_party_means.pdf", plot=gg, device="pdf", 
       width=7, height=4)
gg
```

### Senate
```{r echo=FALSE}
gg <- polarized_plot("Senate")
ggsave("voteview_senate_party_means.png", plot=gg, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_senate_party_means.pdf", plot=gg, device="pdf", 
       width=7, height=4)
gg

```

## Liberal-conservative partisan polarization by chamber
```{r echo=FALSE}
gg <- ggplot(polar_dat_long %>% 
              filter(score == "party.mean.diff.d1"),
              aes(x=year,y=value,group=chamber,col=chamber)) +
              scale_x_continuous(expand = c(0.10, 0), breaks=seq(1880,max(polar_dat_long$year),by=8)) +
              geom_line() + geom_point(size=0.7) + 
              xlab("Year") + ylab("Distance between party means") + 
              geom_text_repel(data=polar_dat_long %>%
                    ungroup() %>%
                    filter(year == min(year) & score=="party.mean.diff.d1"),
                  aes(label = chamber, color = chamber),
                  size = 3,
                  nudge_x = -10,
                  point.padding = 0.1,
                  segment.color = NA,
                  show.legend = FALSE) +
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              scale_color_manual(values=c("House"="#c70828","Senate"="#1372ad"),guide="none") +
              theme_bw() 

ggsave("voteview_party_mean_diff.png", plot=gg, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_party_mean_diff.pdf", plot=gg, device="pdf", 
       width=7, height=4)
gg

```











# Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
