---
title: Polarization in Congress
author: Jeff Lewis
description: Provides updated versions the classic Voteview plots of party mean DW-NOMINATE scores of overtime.  Figures and the underlying data are available for download. 
original_date: "2018-02-01" # Set the original date of your contribution in YYYY-MM-DD format here.
date: "`r format(Sys.time(), '%B %d, %Y')`" # Do not modify this
update_delta: 31
tags:
  - blog # Valid tags are: blog, help, data. You may use multiple tags for a post
  - data
---

```{r echo=FALSE, message=FALSE, error=FALSE}
# We use pacman to load and install dependencies.
library(pacman)

# Please list your article's dependencies here, comma separated:
p_load(tidyverse, ggplot2, ggrepel)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Here we create the polarization data and key graphics the Keith shows and describes in greater detail [here
](https://legacy.voteview.com/political_polarization_2015.htm).  You can download the `Rmarkdown` code that produces this file and all of the data sets and graphics presented [here](party_polarization.Rmd).

This polarization data set includes the following fields (again following Keith):

```
 0.  Chamber
 1.  Congress Number
 2.  First Year of the Congress
 3.  Difference in Party Means - first dimension
 4.  Proportion Moderates
 5.  Proportion of moderate Democrats (-0.25 to +0.25)
 6.  Proportion of moderate Republicans (-0.25 to +0.25)
 7.  Overlap
 8.  Chamber Mean - first dimension
 9.  Chamber Mean - second dimension
10.  Democratic Party Mean - first dimension
11.  Democratic Party Mean - second dimension
12.  Republican Party Mean - first dimension
13.  Republican Party Mean - second dimension
18.  Armed Services Republicans - first dimension
20.  Armed Services Democrats - first dimension
```
# Load data from the polarization csv I file I previously created using the procedures from https://voteview.com/articles/party_polarization
```{r}
 
 polar_dat <- read_csv ("voteview_polarization_data.csv")
 
```

Here are the first few lines of the resulting dataset:

```{r}
head(polar_dat)
write_csv(polar_dat,path="voteview_polarization_data.csv")
```

## Party means on the liberal-conservative dimension over time by chamber

Here are the party means over time plots for each chamber:
```{r echo=FALSE}
polar_dat_long <- polar_dat %>% gather(score,value,-chamber,-year,-congress)
labels <- c("dem.mean.d1"="All D",
            "rep.mean.d1"="All R",
            "ascr"="Armed Svcs R",
            "ascd"="Armed Svcs D")

polarized_plot <- function(chamb) {
  pdatl <- polar_dat_long %>% 
                filter(chamber==chamb,
                       score %in% c("dem.mean.d1","rep.mean.d1",
                                    "north.dem.mean.d1","south.dem.mean.d1")) %>%
                mutate(party=labels[score]) %>%
                ungroup()
  
  gg <- ggplot(data=pdatl,
               aes(x=year,y=value,group=party,col=party)) +
               scale_x_continuous(expand = c(0.15, 0), 
                                 breaks=seq(1991, max(pdatl$year), by=8)) +
               geom_line() + geom_point(size=0.7) + 
               xlab("Year") + ylab("Liberal / Conservative (econ)") + 
               geom_text_repel(data=pdatl %>%
                    filter(year == min(year)),
                  aes(label = party, color = party),
                  size = 3,
                  nudge_x = -8,
                  point.padding = 0.1,
                  segment.color = NA,
                  show.legend = FALSE) +
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              scale_color_manual(values=c("All R"="red","All D"="blue",
                                         "Armed Svcs R"="violet",
                                         "Armed Svcs D"="purple"),guide="none") +
              theme_bw()
  gg
}
```

### House
```{r echo=FALSE}
gg <- polarized_plot("House")
ggsave("voteview_house_party_means.png", plot=gg, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_house_party_means.pdf", plot=gg, device="pdf", 
       width=7, height=4)
gg
```

### Senate
```{r echo=FALSE}
gg <- polarized_plot("Senate")
ggsave("voteview_senate_party_means.png", plot=gg, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_senate_party_means.pdf", plot=gg, device="pdf", 
       width=7, height=4)
gg

```

## Liberal-conservative partisan polarization by chamber
```{r echo=FALSE}
gg <- ggplot(polar_dat_long %>% 
              filter(score == "party.mean.diff.d1"),
              aes(x=year,y=value,group=chamber,col=chamber)) +
              scale_x_continuous(expand = c(0.10, 0), breaks=seq(1880,max(polar_dat_long$year),by=8)) +
              geom_line() + geom_point(size=0.7) + 
              xlab("Year") + ylab("Distance between party means") + 
              geom_text_repel(data=polar_dat_long %>%
                    ungroup() %>%
                    filter(year == min(year) & score=="party.mean.diff.d1"),
                  aes(label = chamber, color = chamber),
                  size = 3,
                  nudge_x = -10,
                  point.padding = 0.1,
                  segment.color = NA,
                  show.legend = FALSE) +
              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
              scale_color_manual(values=c("House"="#c70828","Senate"="#1372ad"),guide="none") +
              theme_bw() 

ggsave("voteview_party_mean_diff.png", plot=gg, device="png", 
       width=7, height=4, dpi=300)
ggsave("voteview_party_mean_diff.pdf", plot=gg, device="pdf", 
       width=7, height=4)
gg

```